### 2015-03-28 Paper Seminar/classification.tex
Added file itself

A wonderful classification. To be edited. Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!