### 2015-03-28 Paper Seminar/classification.tex
Fixed file

While using computer aided systems to create treatment plans, this involves time-consuming calculations during optimizations (as seen in this project). Therefore, a classification attempt is proposed, using prior knowledge to reduce the amount of time taken to process.

\subsection{General Idea}
When classifying, you search for similarities or analogies in your given set of data. These provide a basis for a comparison among previously created optimiziations